### 2013-08-28 doc/pgas13/flat-combining.tex
improve explanation of linearizability

To preserve linearizability for flat-combined operations, they must obey a consistent global order.
First of all, the execution of local combined operations must be serializable;
this is unchanged from shared-memory flat-combining, and is trivially true due to the atomicity enabled by cooperative multithreading.
When combined operations are applied globally, they are serialized at a particular core that owns the synchronization (the \emph{master} core for the Stack).
Whenever a global commit starts, a fresh combiner must be created for subsequent operations to use. If they were to use the old combiner's state to satisfy requests locally, it would violate global ordering because the local object would not reflect other cores' updates.
% This guarantees, for example, that destructive operations only happen once (i.e., a worker cannot locally pop the same value that another core's combined operation does).
% haven't discussed queue, set, or map yet...
%(the ``master'' core for the Queue and Stack, or the corresponding hash cell for the Set and Map). The global order, then, is essentially the concatenation of the cores' sequential orders.
For this serial ``concatenation'' of serialized batches to be valid, the order observed by workers during local combining must be the same as what can be observed globally as operations are committed.

@@

Combining set and map operations expose more nuances in the consistency guarantees.
Insert and lookup operations performed by different tasks are inherently unordered unless synchronized externally.
Therefore, a batch of these operations can be committed globally in parallel, since they are guaranteed not to conflict with each other.
Note that if an insert finds that its key is already in the combiner object, it does not need to send its own message. However, it must still block until that insert is done to ensure that a subsequent operation it performs cannot be reordered with it, preserving program order.
% providing the same guarantee as all other delegate operations. 
% This ensures that if one wanted to use the insert to signal synchronization, it is guaranteed to have been committed globally before the task continues.

Similarly, lookups can piggy-back on other lookups of the same key.
Using intuition from store/write buffers in modern processors, it is tempting to allow a lookup to be satisfied locally by inspecting outstanding inserts.
However, this allows the local order in which keys were inserted to be observed.
If this is allowed, then to preserve SC, that same order must be respected globally, forcing each batch to commit atomically with respect to other cores' batches. Enforcing this is prohibitively expensive, so a cheaper solution is chosen: lookups must get their values from the global data structure, so batches can be performed in parallel.

@@

As mentioned before, the Grappa FC framework takes a different approach than the original flat-combining work for expressing how operations combine.
% As mentioned before, the local \emph{publication list} is trivial to access atomically in Grappa's cooperative scheduling.
% Instead, in the Grappa FC Framework, 
For each global structure instance, a \emph{proxy} object is created on each core and each worker merges its request into that structure before blocking or becoming the combiner.
Each global data structure must define a \emph{proxy} that has \emph{state} to track combined requests, \emph{methods} to add new requests, and a way to \emph{sync} the state globally.
% \begin{enumerate}
%   \item A local \emph{proxy} data structure for tracking combined requests.
%   \item \emph{Combining methods} that operate on the local proxy.
%   \item A \emph{sync} method that globally commits the proxy's state.
% \end{enumerate}
An example proxy declaration for the GlobalStack is shown in Figure~\ref{fig:push}.

>>>>

To behave in accordance with sequential consistency, operations on a particular data structure must obey a consistent global order.
One way to provide this is to guarantee \emph{linearizability}~\cite{herlihy1990linearizability} of operations, which requires that the operation appear to take effect globally at some instantaneous point during invocation of the API call. This ensures that a consistent total order can be imposed on operations on a single data structure.
% To preserve linearizability for flat-combined operations, they must obey a consistent global order.
For operations to be globally linearizable, first of all the execution of local combined operations must be serializable;
this is unchanged from shared-memory flat-combining, and is trivially true due to the atomicity enabled by cooperative multithreading.
Second, combined operations must be committed atomically in some globally serializable order.
When committing, combined operations are serialized at the particular core that owns the synchronization (the \emph{master} core for the Stack).
Whenever a global commit starts, a fresh combiner must be created for subsequent operations to use. If they were to use the old combiner's state to satisfy requests locally, it would violate global ordering because the local object would not reflect other cores' updates.
% This guarantees, for example, that destructive operations only happen once (i.e., a worker cannot locally pop the same value that another core's combined operation does).
% haven't discussed queue, set, or map yet...
%(the ``master'' core for the Queue and Stack, or the corresponding hash cell for the Set and Map). The global order, then, is essentially the concatenation of the cores' sequential orders.
For this serial ``concatenation'' of serialized batches to be valid, the order observed by workers during local combining must be the same as what can be observed globally as operations are committed.

@@

Combining set and map operations exposes more nuances in the requirements for consistency.
Insert and lookup operations performed by different tasks are inherently unordered unless synchronized externally.
Therefore, a batch of these operations can be committed globally in parallel, since they are guaranteed not to conflict with each other.
Note that if an insert finds that its key is already in the combiner object, it does not need to send its own message. However, it must still block until that insert is done to ensure that a subsequent operation it performs cannot be reordered with it, preserving program order.
% providing the same guarantee as all other delegate operations. 
% This ensures that if one wanted to use the insert to signal synchronization, it is guaranteed to have been committed globally before the task continues.
Similarly, lookups can piggy-back on other lookups of the same key.

Using intuition from store/write buffers in modern processors, it is tempting to allow a lookup to be satisfied locally by inspecting outstanding inserts.
However, this would allow the local order in which keys were inserted to be observed.
Then to preserve SC, that same order would need to be respected globally, forcing each batch to commit atomically with respect to other cores' batches. Enforcing this would be prohibitively expensive, so a cheaper solution is chosen: lookups only get their values from the global data structure, therefore batches can be performed in parallel.

@@

As mentioned before, the Grappa FC framework takes a different approach than the original flat-combining work for expressing how operations combine.
% As mentioned before, the local \emph{publication list} is trivial to access atomically in Grappa's cooperative scheduling.
% Instead, in the Grappa FC Framework, definition
For each global structure instance, a \emph{proxy} object is created on each core and each worker merges its request into that structure before blocking or becoming the combiner.
Each global data structure must define a \emph{proxy} that has \emph{state} to track combined requests, \emph{methods} to add new requests, and a way to \emph{sync} the state globally.
% \begin{enumerate}
%   \item A local \emph{proxy} data structure for tracking combined requests.
%   \item \emph{Combining methods} that operate on the local proxy.
%   \item A \emph{sync} method that globally commits the proxy's state.
% \end{enumerate}
An example proxy declaration for the GlobalStack is shown in Figure~\ref{fig:push}.

### 2013-08-28 doc/pgas13/intro.tex
improve explanation of linearizability

% Globally shared data structures are one of the cornerstones of shared-memory and PGAS abstractions. 
% In order for multiple concurrent threads to interact with one another correctly, they must observe shared data consistently.
It is commonly accepted that the easiest consistency model to reason about is \emph{sequential consistency} (SC), which enforces that all accesses are visible in program order and appear to happen in some global serializable order.
To preserve SC, operations on shared data structures should be \emph{linearizable}~\cite{herlihy1990linearizability}; that is, appear to happen atomically in some global total order.
In both physically shared memory and PGAS, maintaining linearizability presents performance challenges.
The simplest way is to have a single global lock to enforce atomicity and linearizability through simple mutual exclusion. Literally serializing accesses in this way is typically considered prohibitively expensive, even in physically shared memory.
% Much prior work has explored the scalability challenges of adding more concurrent accessors to shared data structures in shared memory systems.
However, as the number of concurrent accessors increases, the more they contend and the more synchronization operations fail, even for fine-grained lock-free synchronization schemes.
With the massive amount of parallelism in a cluster of multiprocessors and with the increased cost of remote synchronization, the problem is magnified.

>>>>

% Globally shared data structures are one of the cornerstones of shared-memory and PGAS abstractions. 
% In order for multiple concurrent threads to interact with one another correctly, they must observe shared data consistently.
It is commonly accepted that the easiest consistency model to reason about is \emph{sequential consistency} (SC), which enforces that all accesses are visible in program order and appear to happen in some global serializable order.
To preserve SC, operations on shared data structures should be \emph{linearizable}~\cite{herlihy1990linearizability}; that is, appear to happen atomically in some global total order.
In both physically shared memory and PGAS, maintaining linearizability presents performance challenges.
The simplest way is to have a single global lock to enforce atomicity and linearizability through simple mutual exclusion. Literally serializing accesses in this way is typically considered prohibitively expensive, even in physically shared memory.
% Much prior work has explored the scalability challenges of adding more concurrent accessors to shared data structures in shared memory systems.
However, even in fine-grained lock-free synchronization schemes, as the number of concurrent accessors increases, there is more contention, resulting in more failed synchronization operations.
With the massive amount of parallelism in a cluster of multiprocessors and with the increased cost of remote synchronization, the problem is magnified.
